---
title: "Data Analysis in Social Science  (1 February)"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
dfStable <- read_tsv("/Users/chelsealee/Desktop/SSI3003/UndSoc/tab/ukhls_wx/xwavedat.tab")
```

This is a cross-wave data file with stable characteristics of individuals. See the codebook at https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/search/datafile/xwavedat.

Exercises.

1. Select the variables for: sex (derived), date of birth (doby_dv), ethnic group (racel_dv). Also keep the cross-wave identifier (pidp) and the sample origin variable (memorig).

```{r}
dfStable1 <- dfStable%>%
      select(sex_dv, doby_dv, racel_dv, pidp, memorig)
```


2. Filter the data to keep:
        a) men only.
        b) people born before 1950 and after 1975.
        c) men of Pakistani origin born in 1958 or 1982. 
```{r}
dfStable1%>%
  filter(sex_dv==1)

dfStable1%>%
  filter(doby_dv>1950 | doby_dv>1975)

dfStable1%>%
  filter(sex_dv==1 & (doby_dv==1958|doby_dv==1982) & racel_dv==10)
```


3. Recode birth year into cohorts (a new variable): the G.I. Generation (born before 1924), Silent Generation (1925-42), Baby Boomers (1943-65), Generation X (1966-1980), Millenials (1981-99), Generation Z (2000-). (The years are approximate.)

```{r}
dfStable1 <- dfStable1%>%
  mutate(cohort = case_when(
    doby_dv<1924 ~ "G.I. Generation",
    between(doby_dv, 1925, 1942) ~ "Baby Boomer",
    between(doby_dv, 1966, 1980) ~ "Generation X",
    between(doby_dv, 1981,1999) ~ "Millenials", #doby_dv %in% 1981:1999
    doby_dv>2000 ~ "Generation Z"))
  

```


4. Recode ethnicity into the following groups: white British, Other White, Indian, Pakistani, other. (This classification doesn't make much sense, but we're doing this as an exercise).

```{r}
dfStable1 <- dfStable1 %>%
  mutate(race=recode(racel_dv,
                     "1" = "White British",
                     "2" = "Other White",
                     "3" = "Other White",
                     "4" = "Other White",
                     "9" = "Indian",
                     "10" = "Pakistani",
                     .default = "Other"
  ))
```


5. Count the number of people belonging to different ethnic groups (and produce percentages). 
```{r}
dfStable1 %>%
  count(race)%>%
  mutate(perc = n / sum(n)*100)
```

6. Summarise the proportion of white British by generation.
```{r}
dfStable1 %>%
  filter(race=="White British")%>%
  group_by(cohort)
  
```

