---
title: "week5_lab"
author: "Chelsea Lee"
date: "08/02/2021"
output:
  md_document:
    variant: markdown_github
---
```{r}
library(tidyverse)
library(vroom)
```

Exercises.

1. Calculate the average age of parents (separately for women and men) of children born between waves 9 and 10.

```{r}
# Read the data for newborn children (born between waves 9 and 10).
# Observations are at the parent level.
newborn10 <- vroom("/Users/chelsealee/Desktop/SSI3003/UndSoc/tab/ukhls_w10/j_newborn.tab")
```

```{r}
# Read the data with people's age.

indall10 <- vroom("/Users/chelsealee/Desktop/SSI3003/UndSoc/tab/ukhls_w10/j_indall.tab")

```

```{r}
# use semi-join (filtering join) to keep only new parents in the dataset
indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  # split by gender
  group_by(j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
meanAge = mean(j_age_dv, na.rm = TRUE)
)
```


2. Split the table by ethnic group for the following groups: White British, Indian, Pakistani, Bangladeshi, and African.

```{r}
# ethnicity is recoded in the cross-wave file
stable <- vroom("/Users/chelsealee/Desktop/SSI3003/UndSoc/tab/ukhls_wx/xwavedat.tab",
            col_select = c(pidp, racel_dv))

```

```{r}
indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  # join with ethnicity data
  left_join(stable, by = "pidp") %>%
  # keep only some ethnic groups
  filter(racel_dv %in% c(1, 9, 10, 11, 15)) %>%
  # group by ethnicity and sex
  group_by(racel_dv, j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
meanAge = mean(j_age_dv, na.rm = TRUE)
  )
```

```{r}
indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  filter(j_ethn_dv %in% c(1, 9, 10, 11, 15)) %>%
  # group by ethnicity and sex
  group_by(j_ethn_dv, j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
meanAge = mean(j_age_dv, na.rm = TRUE)
  )
```


3. Produce a data table that includes only twins born between waves 9 and 10. How does the age of their parents compare with the age of parents of non-twin children?

```{r}
twins <- newborn10 %>%
  # find parents with more than one new child
  group_by(pidp) %>%
  # create a variable with the number of children per parent
  mutate(n_children = n()) %>%
  # keep only those with > 1 children
  filter(n_children > 1) %>%
  select(pidp, n_children)

# repeat as in ex.1 but only for the parents of twins.

indall10 %>%
  semi_join(twins, by = "pidp") %>%
  # split by gender
  group_by(j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
meanAge = mean(j_age_dv, na.rm = TRUE)
  )
```




